%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Global Settings for Workflow Parametrization}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
    \frametitle{Outline}
    \begin{columns}[t]
        \begin{column}{.5\textwidth}
            \tableofcontents[sections={1-7},currentsection]
        \end{column}
        \begin{column}{.5\textwidth}
            \tableofcontents[sections={8-15},currentsection]
        \end{column}
    \end{columns}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
  \frametitle{What is this about?}
  \begin{question}[Questions]
   	\begin{itemize}
      \item Avoiding I/O contention! How?
      \item Accounting for FS Latency! How?
    \end{itemize}
  \end{question}
   \begin{docs}[Objectives]
   	 \begin{enumerate} 
        \item Learn how to tune \Snakemake{} to mitigate I/O contention.
        \item Learn how to configure \Snakemake{} to allow for file system latency.
    \end{enumerate}
  \end{docs}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
	\frametitle{\Interlude{Random Access and File System Latency}}
	\begin{hint}
		Here, we explain to users what random access patterns are and how they arise.\newline
		We tell -- roughly -- the story of file system latencies, their causes and characteristics of different file systems and speeds.\newline
		\pause
		We do not feel it's appropriate to lecture you. Instead, we show how easy \Snakemake will offer a solution to I/O contention.
	\end{hint}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Global Workflow Configuration}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]
  \frametitle{\Snakemake{} Profiles}
  \begin{hint}
  	Profiles can shorten your command lines and be an easy remedy for the described issues!
  \end{hint}
  \pause
  Two kinds of profiles are supported:
  \begin{itemize}[<+->]
  	\item A global profile that is defined in a system-wide or user-specific configuration directory (on Linux, this will be \altverb{\~/.config/snakemake} and \altverb{/etc/xdg/snakemake}, you can find the answer for your system via \altverb{snakemake --help}).
  	\item A workflow specific profile that is defined via a flag (\altverb{--workflow-profile}) or searched in a default location (profile/default) in the working directory or next to the \altverb{Snakefile}.
  \end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]
  \frametitle{Your Profile}
    \begin{onlyenv}<1| handout:0>
      Our first line forces the use of the \lhref{https://yte-template-engine.github.io/}{\bf{Y}AML \bf{T}emplate \bf{E}ngine} to parse the following lines accordingly.
      \begin{lstlisting}[language=Bash, style=Shell]
__use_yte__: true  	 
      \end{lstlisting}
    \end{onlyenv}
    \begin{onlyenv}<2| handout:0>
      Now, our default executor is SLURM.
      \begin{lstlisting}[language=Bash, style=Shell]
__use_yte__: true
executor: slurm
     \end{lstlisting}
     No more, \altverb{snakemake --executor slurm ...}!
   \end{onlyenv}
   \begin{onlyenv}<3| handout:0>
   	 The next line tells Snakemake to wait for a minute, if output files are not present. This is more than enough time, even for NFS-Filesystems. 
   	 \begin{lstlisting}[language=Bash, style=Shell]
__use_yte__: true
executor: slurm
latency-wait: 60
   	 \end{lstlisting}
    \end{onlyenv}
    \begin{onlyenv}<4| handout:0>
      The entire rest, will tell the storage plugin (\altverb{snakemake-storage-plugin-fs}) to stage in to the node-local storage on Mogon, for \emph{every} job and to copy back your results. When dealing with I/O intensive jobs, this can boost your performance tremendously. 
      \begin{lstlisting}[language=Bash, style=Shell]
__use_yte__: true
executor: slurm
latency-wait: 60
default-storage-provider: fs
shared-fs-usage:
  - persistence
  - sources
  - source-cache
local-storage-prefix: /localscratch/$SLURM_JOB_ID
    \end{lstlisting}
   \end{onlyenv}
  \begin{onlyenv}<5| handout:0>
  	%TODO remove this part, once the snakemake release is ready
  	\begin{warning}
  	  Currently, we are working on way to annotate susceptible I/O pattern, the highlighted parts are not yet functional.
  	\end{warning}
  	%You may copy this setup from \texttt{<++course.pathtosetup++>/config.yaml} to the \altverb{\~/.config/snakemake}-folder - unless your local admins provide a cluster-wide configuration.
  	\begin{lstlisting}[language=Bash, style=Shell]
__use_yte__: true
executor: slurm
@latency-wait: 60@
@default-storage-provider: fs@
@shared-fs-usage:@
@  - persistence@
@  - sources@
@  - source-cache@
@local-storage-prefix: /localscratch/$SLURM_JOB_ID@
  	 \end{lstlisting}
  \end{onlyenv}
   \begin{onlyenv}<6| handout:1>
   	The complete configuration out to be in \altverb{\~/.config/snakemake/config.yaml}
   	\begin{lstlisting}[language=Bash, style=Shell]
__use_yte__: true
executor: slurm
latency-wait: 60
default-storage-provider: fs
shared-fs-usage:
- persistence
- sources
- source-cache
local-storage-prefix: /localscratch/$SLURM_JOB_ID
    \end{lstlisting}
  \end{onlyenv}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]
	\frametitle{\HandsOn{Running with a Profile I}}
	Enter 
	\begin{lstlisting}[language=Bash, style=Shell]
export SNAKEMAKE_PROFILE="$HOME/.config/snakemake"
    \end{lstlisting}
	in your \altverb{.bashrc}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]
	\frametitle{\HandsOn{Running with a Profile II}}
	Copy the configuration file:
	\begin{lstlisting}[language=Bash, style=Shell]
$ mkdir -p ~/.config/snakemake
$ cp ~/workflows/config.yaml \
> ~/.config/snakemake/.
	\end{lstlisting}
    Activate the new settings:
    \begin{lstlisting}[language=Bash, style=Shell]
$ source ~/.bashrc
    \end{lstlisting}
    And run the workflow - just for fun:
    \begin{lstlisting}[language=Bash, style=Shell]
$ snakemake -F # just this one flag!
    \end{lstlisting}
\end{frame}

	

