%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Parametizing your Workflow - II}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
    \frametitle{Outline}
    \begin{columns}[t]
        \begin{column}{.5\textwidth}
            \tableofcontents[sections={1-9},currentsection]
        \end{column}
        \begin{column}{.5\textwidth}
            \tableofcontents[sections={10-18},currentsection]
        \end{column}
    \end{columns}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
  \frametitle{What is this about?}
  \begin{question}[Questions]
   	\begin{itemize}
      \item How do we add execution parameters?
      \item How do we tune scientific parameters?
    \end{itemize}
  \end{question}
   \begin{docs}[Objectives]
   	 \begin{enumerate} 
        \item Learn to use parameters relevant for the batch systems.
        \item Learn how to tune \Snakemake{} on the command line.
        \item Learn how to tune \Snakemake{} with configuration files.
    \end{enumerate}
  \end{docs}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Configuration File}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}[fragile]
	\frametitle{The \texttt{Snakemake} \texttt{resources} Section}
	\texttt{Snakemake} rules provide an additional \altverb{resource} section:
	\begin{lstlisting}[language=Python,style=Python]
rule <name>:
	...
	resources:
		partition='parallel',
		mem_mb=1800,
		cpus_per_task=4
	\end{lstlisting}
	\begin{hint}
		Note the \textbf{,}!
	\end{hint}
	\pause
	\begin{docs}
		You \emph{may} define \emph{any} resource keyword within any rule.
	\end{docs}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
	\frametitle{The \texttt{Snakemake} \texttt{resources} Section - its Downside}
	\begin{block}{Every Resource Spec needs a Change per Rule???}
		You might have noticed, this specification per rule is most untidy. \texttt{Snakemake}'s design principle is: ship workflows which run \emph{everywhere} \& \emph{every time}.
		\newline \pause
		Relax: Every parameter can be specified:
		\begin{itemize}
			\item in a \texttt{Snakefile},
			\item on the command line,
			\item and re-usable in configuration files!
		\end{itemize}
	\end{block}
\end{frame} 