%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Getting your Sample Data}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
	\frametitle{Outline}
	\begin{columns}[t]
		\begin{column}{.5\textwidth}
			\tableofcontents[sections={1-9},currentsection]
		\end{column}
		\begin{column}{.5\textwidth}
			\tableofcontents[sections={10-18},currentsection]
		\end{column}
	\end{columns}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
	\frametitle{What is this about?}
	\begin{question}[Questions]\begin{itemize}
			\item No Question: We just need to get sample data!
		\end{itemize}
	\end{question}
	\begin{docs}[Objectives]
		\begin{enumerate}
			\item And that, getting sample data and software definitions is our objective.
		\end{enumerate}
	\end{docs}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{frame}[fragile]
  \frametitle{\HandsOn{Getting Your Course Material I}}
  We shall copy a few install scripts (which also will download some sample data).\newline
  Please copy the directory "\texttt{\configparam{pathtosetup}}".\newline
  \begin{hint}
  	Remember, to copy an entire directory, you can use:
  	\begin{lstlisting}[language=Bash, style=Shell]
$ cp -r <path> ~/.
  	\end{lstlisting}
    We assume that you are working in your \altverb{HOME}. Then typing the \textasciitilde{} is not necessary.
  \end{hint}
  \pause
  Now, change into this directory!  
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{frame}[fragile]
  \frametitle{\HandsOn{Getting Your Course Material II}}
  \begin{columns}[t]
	 \begin{column}{.5\textwidth}
		When copied, \altverb{cd} into this directory.\newline The command \altverb{tree} should show:
	 \end{column}
	 \begin{column}{.5\textwidth}
	   \begin{minipage}[t]{0.5\textwidth}
		 \setstretch{0.1}
		 {\footnotesize \DTsetlength{0.2em}{1em}{0.2em}{0.4pt}{.6pt}
			 \dirtree{%
                 .1 {.}.
                 .2 {condarc}.
                 .2 {get\_tutorial.sh}.
                 .2 {tutorial}.
                 .3 {01\_Snakefile}.
                 .3 {02\_Snakefile}.
                 .3 {\ldots}.
			}}
      \end{minipage}
	\end{column}
  \end{columns}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{frame}[fragile]
  \frametitle{\HandsOn{Obtaining Your Tutorial Sample Data}}
  Please run the \altverb{get_data.sh} script you just copied:
  \begin{lstlisting}[language=Bash, style=Shell,basicstyle=\footnotesize]
$ bash get_data.sh
  \end{lstlisting}
  \begin{task}
  	This script will download and unpack the sample data for this course. Please take a look in this script to understand it. Where are your sample data after running this script?
  \end{task}
\end{frame}